---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r, warning=FALSE, message=FALSE}

library(tidyverse)

av %>% 
  select(
    Name.Alias,
    starts_with("Death")
  ) %>% 
  head()


deaths <- av %>% 
  pivot_longer(
    starts_with("Death"),
    names_to = "Time",
    values_to = "Died"
  ) %>% 
  select(
    URL, Name.Alias, Time, Died
  )
head(deaths)
View(deaths)


```

Similarly, deal with the returns of characters.

```{r, message=FALSE, warning=FALSE}

av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)

av %>% 
  select(
    Name.Alias,
    starts_with("Return")
  ) %>% 
  head()

returns <- av %>% 
  pivot_longer(
    starts_with("Return"),
    names_to = "Time",
    values_to = "Returned"
  ) %>% 
  select(
    URL, Name.Alias, Time, Returned
  )
head(returns)
View(returns)


```

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r, message=FALSE, warning=FALSE}

numdeaths <- deaths %>%
  filter(Died == "YES") %>%
  count(URL) %>% summarise(averagedeaths = sum(n)/nrow(av))
  #had to use URL because some of the names were null and then it made one row with 7 deaths for all null names

#average number of deaths per avenger is .5144509
numdeaths$averagedeaths

```
  
Average number of deaths for an avenger is .5144509  


## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.  

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

  
### - Kaitlyn Hoyme -  


> "But you can only tempt death so many times. There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death. " - fivethirtyeight.com  



```{r, message=FALSE, warning=FALSE}

deadavengers <- av %>%
  filter(Death1 == "YES") %>%
  count(URL) 

returnedavengers <- av %>%
  filter(Death1 == "YES", Return1 == "YES") %>%
  count(URL) 

#count(returnedavengers)$n
#46 avengers died and came back once
#count(deadavengers)$n
#69 avengers died once

firstdeathreturn <- count(returnedavengers)$n / count(deadavengers)$n
firstdeathreturn

#This is 2/3, so the first part of the statement is correct, as we can see after their first death, 2/3 of the avengers came back.

diedtwice <- av %>% 
  filter(Death2=="YES") %>% count(URL)
#count(diedtwice)$n
#16 avengers died twice

diedthrice <- av %>% 
  filter(Death3=="YES") %>% count(URL)
#count(diedthrice)$n
#2 avengers died 3 times

returnedafter2 <- av %>%
   filter(Return2 == "YES") %>% count(URL)
#returnedafter2

returnedafter3 <- av %>%
    filter(Return3 == "YES") %>% count(URL)
#returnedafter3

#count(returnedafter2)$n
#8 avengers returned after their second death


#return after dying twice
count(returnedafter2)$n / count(diedtwice)$n

#return after dying thrice
count(returnedafter3)$n / count(diedthrice)$n

#both of these = .5, so we see the statement is true.

```


As we can confirm by looking at the code, the statement was correct in saying that after the first death there is about a 2/3 chance of returning, and after the 2nd or 3rd there is a 50% chance.  

### - Cassidy Berghoff -  

> "The MVP of the Earth-616 Marvel Universe Avengers has to be Jocasta — an android based on Janet van Dyne and built by Ultron — who has been destroyed five times and then recovered five times." - fivethirtyeight.com

```{r, message=FALSE, warning=FALSE}

numDeath = deaths %>% filter(Name.Alias == "Jocasta", Died == "YES") %>% nrow()
numReturn = returns %>% filter(Name.Alias == "Jocasta", Returned == "YES") %>% nrow()

sprintf("The number of times Jocasta has died is: %d. The number of times Jocasta has returned is: %d.", numDeath, numReturn)



```
As seen above, the statement is correct. Jocasta specifically has died 5 times and has returned all 5 times.


### - Team member -  

### - Team member -  
